\textit{Supplementary materials for:} Fitness tracking reveals task-specific associations between memory, mental health, and exercise

### 2021-10-13 paper/main.tex
added exploratory correlation analysis

\newcommand{\devices}{S1}
\newcommand{\frDetail}{S2}
\newcommand{\natDetail}{S3}
\newcommand{\vocabDetail}{S4}
\newcommand{\spatialDetail}{S5}
\newcommand{\fitDists}{S6}
\newcommand{\fitDistgridImmediate}{S7}
\newcommand{\fitScatterImmediate}{S8}
\newcommand{\fitDistgridDelayed}{S9}
\newcommand{\fitScatterDelayed}{S10}

@@

\subsubsection*{Exploratory correlation analyses}
We used a bootstrap procedure to identify reliable correlations
between different memory-related, fitness-related, and
demographic-related variables.  For each of $N = 1000$ iterations, we
selected (with replacement) a sample of 113 participants to include.
This yielded, for each iteration, a sampled ``data matrix'' with one
row per sampled participant and one column for each measured variable.
When participants were sampled multiple times in a given iteration, as
was often the case, this matrix contained duplicate rows.  We used a
round-robin imputation procedure to estimate the values of any missing
features~\citep{Buck60}.  Next, we computed the Pearson's correlation
between each pair of columns.  This yielded, for each pair of columns,
a distribution of $N$ bootstrapped correlation coefficients.  If fewer
than $95\%$ of the coefficients for a given pair of columns had the
same sign, we excluded the pair from further analysis and considered
the expected correlation between those columns to be undefined.  If
$\geq 95\%$ of the coefficients for a given pair of columns had the
same sign, we computed the expected correlation coefficient as:
\[
  \mathbb{E}_{i, j}\left[ r\right] = \tanh\left(\frac{1}{N} \sum_{n=1}^N
  \tanh^{-1}(\mathrm{corr}\left(m(i)_n, m(j)_n\right))\right),
\]
where $m(x)_n$ represents column $x$ of the bootstrapped data matrix
for iteration $n$, $\tanh$ is the hyperbolic tangent, and $\tanh^{-1}$
is the inverse hyperbolic tangent.

@@

We used the same bootstrap procedure described above (used in our
exploratory correlation analyses) to generate $N = 1000$ bootstrapped
data matrices whose rows reflected sampled participants and whose
columns reflected different measured variables.

@@

We constructed a ``null'' distribution by using the same procedure as
above, but where the columns in the test datasets were randomly
permuted with each iteration (thereby breaking any meaningful
predictive information between the training and test data).  We
assessed the statistical significance ($p$-values) of the observed
RMSD values by computing the proportions of null RMSD values that were
less than the observed value.  We also assessed the significance of
the observed regression weights using $t$-tests to compare the means
of the observed versus null distributions of weights.

@@

\begin{itemize}
% \item characterizing behaviors (color by quartile and continue the
%   color scheme in later figures-- hue reflects task, shading reflects
%   performance.  white outline means immediate, black outline means delayed)
%   \begin{itemize}
%   \item Free recall (immediate + delayed): pfr, lag-CRP, spc  (color:
%     recall performance).  
%     Figure~\ref{fig:fr_behavioral}.
%     \item Naturalistic recall (immediate + delayed): reproduce a version of the sherlock
%       movie/recall trajectories (color: mean precision).  Figure~\ref{fig:nat_behavioral}.
%     \item Foreign language flashcards (immediate + delayed):
%       p(correct) histogram (color: p(correct)).  Figure~\ref{fig:vocab_behavioral}.
%       \item Spatial learning: mean error by number of shapes (color:
%         intercept and slope of line fit to errors as a function of the number of
%         shapes).  Figure~\ref{fig:spatial_behavioral}.
%       \end{itemize}
%     \item Fitness info (break down by task performance, potentially
%       separately for each task); also separate out recent (raw) and
%       recent versus baseline -- color using same color scheme as
%       behavior figure
%       \begin{itemize}
%       \item activity (steps, zone minutes, floors/elevation)
%       \item resting heart rate
%         \item sleep
%         \end{itemize}
  \item exploratory analysis (correlations).  Possibly make some sort of scatter matrix or pairplot-- rows/columns:
          tasks.  Diagonal: histogram of performance metric for that
          task.  Above-diagonal entries: compare performance across
          tasks (by subject).  Below-diagonal entries: empty?
          density/2d histograms?
    \begin{itemize}
    \item Memory-memory
    \item fitness-fitness
    \item survey-survey
    \item (fitness + survey)-memory
    \end{itemize}
  \item predictive analysis (regressions)
    \begin{itemize}
    \item Predict memory performance on held-out task from other tasks
    \item Predict memory performance on each task using fitness data
      \item Predict memory performance on each task using survey data
      \end{itemize}
    \item Reverse correlations: look at recent changes versus baseline
      trends (color using same scheme as behavior figures).  Possibly 
      \begin{itemize}
      \item Fitness profile that predicts performance on each task (barplots + timelines)
      \item Fitness profile for each survey demographic (barplots + timelines)
        \begin{itemize}
          \item Select out mental health demographics (based on meds, stress levels)
          \end{itemize}
        \end{itemize}
  \end{itemize}

>>>>

\newcommand{\devices}{S1}
\newcommand{\frDetail}{S2}
\newcommand{\natDetail}{S3}
\newcommand{\vocabDetail}{S4}
\newcommand{\spatialDetail}{S5}
\newcommand{\fitDists}{S6}
\newcommand{\fitDistgridImmediate}{S7}
\newcommand{\fitScatterImmediate}{S8}
\newcommand{\fitDistgridDelayed}{S9}
\newcommand{\fitScatterDelayed}{S10}
\newcommand{\behBehCorr}{S11}
\newcommand{\fitFitCorr}{S12}
\newcommand{\demDemCorr}{S13}
\newcommand{\allCorr}{S14}

@@

\subsubsection*{Exploratory correlation analyses}
We used a bootstrap procedure to identify reliable correlations
between different memory-related, fitness-related, and
demographic-related variables.  For each of $N = 1000$ iterations, we
selected (with replacement) a sample of 113 participants to include.
This yielded, for each iteration, a sampled ``data matrix'' with one
row per sampled participant and one column for each measured variable.
When participants were sampled multiple times in a given iteration, as
was often the case, this matrix contained duplicate rows.  Next, we computed the Pearson's correlation
between each pair of columns.  This yielded, for each pair of columns,
a distribution of $N$ bootstrapped correlation coefficients.  If fewer
than $97.5\%$ of the coefficients for a given pair of columns had the
same sign, we excluded the pair from further analysis and considered
the expected correlation between those columns to be undefined.  If
$\geq 97.5\%$ of the coefficients for a given pair of columns had the
same sign (corresponding to a bootstrap-estimated two-tailed $p$
threshold of 0.05), we computed the expected correlation coefficient as:
\[
  \mathbb{E}_{i, j}\left[ r\right] = \tanh\left(\frac{1}{N} \sum_{n=1}^N
  \tanh^{-1}(\mathrm{corr}\left(m(i)_n, m(j)_n\right))\right),
\]
where $m(x)_n$ represents column $x$ of the bootstrapped data matrix
for iteration $n$, $\tanh$ is the hyperbolic tangent, and $\tanh^{-1}$
is the inverse hyperbolic tangent.

@@

We used the same bootstrap procedure described above (used in our
exploratory correlation analyses) to generate $N = 1000$ bootstrapped
data matrices whose rows reflected sampled participants and whose
columns reflected different measured variables.  We used a
round-robin imputation procedure to estimate the values of any missing
features~\citep{Buck60}.  This imputation procedure was applied
independently for input features and output features to prevent data contamination.

@@

We constructed a ``null'' distribution by using the same procedure as
above, but where the columns in the test datasets were randomly
permuted with each of $M = 1000$ iterations (thereby breaking any meaningful
predictive information between the training and test data).  We
assessed the statistical significance ($p$-values) of the observed
RMSD values by computing the proportions of null RMSD values that were
less than the observed value.  We also assessed the significance of
the observed regression weights using $t$-tests to compare the means
of the observed versus null distributions of weights.

@@

We wondered about potential links between the different aspects of
participants' data.  For example, if people who exercised in a
particular way also tended to perform better on a given memory task,
this could suggest that either (a) some property intrinsic to participants who
exercised in a particular way might also affect their memory
performance on the given task, and/or (b) particular exercise behaviors
could have a causal impact on memory performance.  We carried out an
exploratory analysis whereby we used a bootstrap-based approach (see
\textit{Exploratory correlation analyses}) to
identify reliable correlations between different aspects of memory
performance (Fig.~\behBehCorr), different aspects of fitness (Fig.~\fitFitCorr), different demographic
attributes (Fig.~\demDemCorr), and correlations between memory
performance, fitness information, and demographic attributes
(Fig.~\allCorr).  Several patterns emerged.  First, we found that
participants' performance on the (within-task) immediate versus delayed memory tests
from the free recall, naturalistic recall, and foreign language
flashcards tasks were positively correlated ($r$s $> 0.25$, $p$s $< 0.05$,
bootstrap corrected).  This suggests that, within each of these tasks,
similar processes or constraints may influence both short term and
long term information retrieval.  We also found reliable across-task correlations
between participants' (immediate and delayed) performance on the free
recall and foreign language flashcards tasks ($r$s $> 0.3$, $p$s $< 0.05$,
bootstrap corrected).

A large number of fitness-related measures displayed reliable
correlations (for a complete report, see Fig.~\fitFitCorr).  For
example, body mass index (BMI) and weight were correlated
($r = 0.91, p < 0.05$, bootstrap corrected).  Resting heart rate over
the prior week was negatively correlated with recent high-intensity
activity levels ($r = 0.39, p < 0.05$, bootstrap corrected).
Participants' peak heart rate (averaged over the prior week) were also
negative correlated with recent decreases in step counts and daily
elevation gains ($r\mathrm{s} < -0.27, p\mathrm{s} < 0.05$, boostrap
corrected), where recent changes were defined as the average values
over the seven days leading up to the test day divided by the average
values over the preceding 30 days.  Although several demographic
attributes (Fig.~\demDemCorr) displayed trivial correlations (e.g.,
participants identifying as male never reported identifying as female,
and so on), we also observed a negative correlation between reported
stress and alertness ($r = -0.44, p < 0.05$, bootstrap corrected), and
positive correlations between the reported task difficulties for all
tasks ($r\mathrm{s} > 0.26, p\mathrm{s} < 0.05$, bootstrap corrected).

We also found reliable correlations between participants' fitness and
demographic measures and their behaviors in different tasks
(for a complete report, see Fig.~\allCorr).  For example, recent
low-intensity (``fat burn'') cardiovascular activity was positively
correlated with immediate
($r = 0.44, p < 0.05$, bootstrap corrected) and delayed ($r = 0.37, p
< 0.05$, bootstrap corrected) recall on the naturalistic memory task.
Moderate intensity (``cardio'') cardiovascular activitiy was
positively correlated with immediate naturalistic recall ($r = 0.47, p
< 0.05$, bootstrap corrected) and immediate recall on the foreign
language flashcards task ($r = 0.43, p < 0.05$, bootstrap corrected).
Recent low-intensity (``out-of-range'') cardiovascular activity was
negatively correlated with spatial memory performance ($r = -0.31, p <
0.05$, bootstrap corrected) whereas recent high-intensity (``peak'')
cardiovascular activity was positively correlated with spatial memory
performance ($r = 0.34, p < 0.05$, bootstrap corrected).  Recent
increases in high-intensity cardiovascular activity (more activity
over the past 7 days than during the preceding 30 days) also predicted
spatial memory performance ($r = 0.41, p < 0.05$, bootstrap
corrected).  Coffee consumption was (slightly) positively correlated with
immediate free recall performance ($r = 0.17, p < 0.05$, bootstrap
corrected) and (slightly) negatively correlated with immediate
naturalistic recall performance ($r = -0.16, p < 0.05$, bootstrap
corrected).

\begin{itemize}
% \item characterizing behaviors (color by quartile and continue the
%   color scheme in later figures-- hue reflects task, shading reflects
%   performance.  white outline means immediate, black outline means delayed)
%   \begin{itemize}
%   \item Free recall (immediate + delayed): pfr, lag-CRP, spc  (color:
%     recall performance).  
%     Figure~\ref{fig:fr_behavioral}.
%     \item Naturalistic recall (immediate + delayed): reproduce a version of the sherlock
%       movie/recall trajectories (color: mean precision).  Figure~\ref{fig:nat_behavioral}.
%     \item Foreign language flashcards (immediate + delayed):
%       p(correct) histogram (color: p(correct)).  Figure~\ref{fig:vocab_behavioral}.
%       \item Spatial learning: mean error by number of shapes (color:
%         intercept and slope of line fit to errors as a function of the number of
%         shapes).  Figure~\ref{fig:spatial_behavioral}.
%       \end{itemize}
%     \item Fitness info (break down by task performance, potentially
%       separately for each task); also separate out recent (raw) and
%       recent versus baseline -- color using same color scheme as
%       behavior figure
%       \begin{itemize}
%       \item activity (steps, zone minutes, floors/elevation)
%       \item resting heart rate
%         \item sleep
%         \end{itemize}
  % \item exploratory analysis (correlations).  Possibly make some sort of scatter matrix or pairplot-- rows/columns:
  %         tasks.  Diagonal: histogram of performance metric for that
  %         task.  Above-diagonal entries: compare performance across
  %         tasks (by subject).  Below-diagonal entries: empty?
  %         density/2d histograms?
  %   \begin{itemize}
  %   \item Memory-memory
  %   \item fitness-fitness
  %   \item survey-survey
  %   \item (fitness + survey)-memory
  %   \end{itemize}
  \item predictive analysis (regressions)
    \begin{itemize}
    \item Predict memory performance on held-out task from other tasks
    \item Predict memory performance on each task using fitness data
      \item Predict memory performance on each task using survey data
      \end{itemize}
    \item Reverse correlations: look at recent changes versus baseline
      trends (color using same scheme as behavior figures).  Possibly 
      \begin{itemize}
      \item Fitness profile that predicts performance on each task (barplots + timelines)
      \item Fitness profile for each survey demographic (barplots + timelines)
        \begin{itemize}
          \item Select out mental health demographics (based on meds, stress levels)
          \end{itemize}
        \end{itemize}
  \end{itemize}

### 2021-10-13 paper/supplement.tex
updated trajectories, added sleep stats
added exploratory correlation analysis

\documentclass[10pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{geometry}
\usepackage[sort]{natbib}
\usepackage{pxfonts}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{hyperref}
\usepackage{lineno}
\usepackage{authblk}

@@

\begin{figure}[p]
\centering
\includegraphics[width=1\textwidth]{figs/free_recall_behavior}
\caption{\textbf{Free recall behavioral results.}}
\label{fig:fr_behavioral}
\end{figure}

\begin{figure}[p]
\centering
\includegraphics[width=1\textwidth]{figs/naturalistic_recall_behavior}
\caption{\textbf{Naturalistic recall behavioral results.}}
\label{fig:nat_behavioral}
\end{figure}

@@

\begin{figure}[p]
\centering
\includegraphics[width=\textwidth]{figs/fitness_distributions}
\caption{\textbf{Distributions of fitness measures.}}
\label{fig:fitness_dists}
\end{figure}

\begin{figure}[p]
\centering
\includegraphics[width=\textwidth]{figs/fitness_dists_immediate}
\caption{\textbf{Distributions of fitness measures, broken down
    by immediate task performance.}}
\label{fig:fitness_dists_immediate}
\end{figure}

\begin{figure}[p]
\centering
\includegraphics[width=\textwidth]{figs/fitness_scatter_immediate}
\caption{\textbf{Scatterplots of fitness measures versus
    immediate task performance measures.}}
\label{fig:fitness_scatters_immediate}
\end{figure}

\begin{figure}[p]
\centering
\includegraphics[width=\textwidth]{figs/fitness_dists_delayed}
\caption{\textbf{Distributions of fitness measures, broken down
    by delayed task performance.}}
\label{fig:fitness_dists_delayed}
\end{figure}

\begin{figure}[p]
\centering
\includegraphics[width=\textwidth]{figs/fitness_scatter_delayed}
\caption{\textbf{Scatterplots of fitness measures versus
    delayed task performance measures.}}
\label{fig:fitness_scatters_delayed}
\end{figure}

>>>>

\documentclass[10pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{geometry}
\usepackage[sort]{natbib}
\usepackage{pxfonts}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{hyperref}
\usepackage{lineno}
\usepackage{authblk}
\usepackage{pdflscape}
\usepackage{rotating}

@@

\begin{sidewaysfigure}[p]
\centering
\includegraphics[width=1\textwidth]{figs/free_recall_behavior}
\caption{\textbf{Free recall behavioral results.}}
\label{fig:fr_behavioral}
\end{sidewaysfigure}

\begin{sidewaysfigure}[p]
\centering
\includegraphics[width=1\textwidth]{figs/naturalistic_recall_behavior}
\caption{\textbf{Naturalistic recall behavioral results.}}
\label{fig:nat_behavioral}
\end{sidewaysfigure}

@@

\begin{sidewaysfigure}[p]
\centering
\includegraphics[width=\textwidth]{figs/fitness_distributions}
\caption{\textbf{Distributions of fitness measures.}}
\label{fig:fitness_dists}
\end{sidewaysfigure}

\begin{sidewaysfigure}[p]
\centering
\includegraphics[width=\textwidth]{figs/fitness_dists_immediate}
\caption{\textbf{Distributions of fitness measures, broken down
    by immediate task performance.}}
\label{fig:fitness_dists_immediate}
\end{sidewaysfigure}

\begin{sidewaysfigure}[p]
\centering
\includegraphics[width=\textwidth]{figs/fitness_scatter_immediate}
\caption{\textbf{Scatterplots of fitness measures versus
    immediate task performance measures.}}
\label{fig:fitness_scatters_immediate}
\end{sidewaysfigure}

\begin{sidewaysfigure}[p]
\centering
\includegraphics[width=\textwidth]{figs/fitness_dists_delayed}
\caption{\textbf{Distributions of fitness measures, broken down
    by delayed task performance.}}
\label{fig:fitness_dists_delayed}
\end{sidewaysfigure}

\begin{sidewaysfigure}[p]
\centering
\includegraphics[width=\textwidth]{figs/fitness_scatter_delayed}
\caption{\textbf{Scatterplots of fitness measures versus
    delayed task performance measures.}}
\label{fig:fitness_scatters_delayed}
\end{sidewaysfigure}

\begin{figure}[p]
\centering
\includegraphics[width=0.4\textwidth]{figs/behavior_behavior_correlations}
\caption{\textbf{Bootstrap-estimated reliable correlations between
    behavioral measures.}}
\label{fig:behavioral_corrs}
\end{figure}

\begin{figure}[p]
\centering
\includegraphics[width=\textwidth]{figs/fitness_fitness_correlations}
\caption{\textbf{Bootstrap-estimated reliable correlations between
    fitness measures.}}
\label{fig:fitness_corrs}
\end{figure}

\begin{figure}[p]
\centering
\includegraphics[width=\textwidth]{figs/survey_survey_correlations}
\caption{\textbf{Bootstrap-estimated reliable correlations between
    demographic measures.}}
\label{fig:survey_corrs}
\end{figure}

\begin{sidewaysfigure}[p]
\centering
\includegraphics[width=\textwidth]{figs/behavior_fitness+survey_correlations}
\caption{\textbf{Bootstrap-estimated reliable correlations between
    behavioral measures and fitness or demographic measures.}}
\label{fig:survey_corrs}
\end{sidewaysfigure}
